---
title: "Motor Trend Magzine - Regreesion Models"
author: "Syed Shaheryar Tirmizi"
date: "8/28/2020"
output: 
html_document: default
---

## Executive Summary

The basic aim of this assignment is to analyze which transmission mode (Auto or Manual) is better in terms of Miles per gallon.
All I have to do is to use linear regression model that will minimize the error and has required variables.
mtcars dataset, provided by r, is used in this analysis.
The analysis also focuses on relationship of mpg with other variables.

## 1. Preliminary Analysis

```{r}
library(datasets)
library(ggplot2)
data("mtcars")
dim(mtcars)
str(mtcars)
summary(mtcars)
head(mtcars)
```
So, the dataset contains 11 variables and 32 records.
Moreover, am variable shows transmission mode (1 for manual & 0 for automatic)

## 2. Data Pre-processing

```{r}
mtcars$am<-as.factor(mtcars$am)
levels(mtcars$am)<-c("Automatic", "Manual")
```
The am variable is now a factor of 2.
Let's check it now

```{r}
class(mtcars$am)
head(mtcars)
```

## 3. Hypothesis Validation
Let's just consider 2 variables only i-e mpg and am.
Null Hypothesis states that there is no difference between mean of trasnmission modes.
In other words, it states that transmission modes has no effect on mgp.
Let's validate it with t test.

```{r}
t.test(mtcars$mpg~mtcars$am)
```

Since P value (0.001) is less then the Threshold value (0.05), hence we can reject the null hypothesis.
It means that there is difference between means of automatic and manual mode (Alternative Hypothesis)

## 4. Regression Analysis

To further investigate Alternative hypothesis, let's do some regression modeling.

```{r}
model1<-lm(mpg~am,mtcars)
summary(model1)
```
By this simple model, we can conclude that, on average, Manual mode cars travel 7.245 miles per gallon more than that of Automatic mode cars provided that other variables remain constant.
This model is not efficient as Adjusted R Squared value (0.3345) suggests that it accounts or only 33.45% variance.

Let's see what happens if we add all variables in our model

```{r}
model2<-lm(mpg~., mtcars)
summary(model2)
```
This model is good and accounts for 80.66% variance.

To choose best model, step function in r is useful that returns the best model acoording to the given parameters. in my case, I didn't add any other parameters and used the default setting.

```{r}
bestModl<-step(model2, trace = 0)
summary(bestModl)
```
Now this model accounts for 83.36% variance and includes 3 variables i-e Transmission mode (am), weight (wt) and qsec (1/4 mile time).

## 5. Conclusion

1. From simple regression model, it is concluded that Manual mode is better than Automatic mode with other variable remain constant

2. negative sign with wt indicates that weight is negatively associated with mpg.
Moreover, this model suggests that Manual mode is better than Auto mode by 2.9358 miles.

## Apendix

### Appendix 1: Boxplot of mpg with Transmission mode

```{r}
g<-ggplot(mtcars, aes(x=am, y=mpg))+geom_boxplot(outlier.colour = "black", outlier.shape = 2, fill="purple")
g<-g + labs(title = "Comparative effect of Automatic and manual cars on Miles per Gallon (mpg)", x="Transimission mode", y="Miles per gallon")
g
```

### Appendix 2: Residual Plots

```{r}
par(mfrow=c(2,2)) 
plot(bestModl)
```