---
title: "Tooth Growth Dataset - Statistical Analysis"
author: "Syed Shaheryar Tirmizi"
date: "7/27/2020"
output:
  html_document: default
---
## Overview
### In this exercise ,I am going to analyze the Toothgrowth dataset provided in the r package datasets. The dataset contains the data of growth of teeth by giving Vitamin c (VC) and orange juice (OJ). It also has 3 dosage quantity i-e 0.5mg, 1mg, 2mg.

## 1. Loading the dataset and neccessary libraries

```{r}
library(ggplot2)
data(ToothGrowth)
```

## 2. Basic summary and struture of dataset

```{r}
head(ToothGrowth)
str(ToothGrowth)
summary(ToothGrowth)
dim(ToothGrowth)
```

## 3. Basic exploratory Analysis

```{r}
g<- ggplot(data=ToothGrowth, aes(x=paste(supp,dose) , y=len, fill=factor(dose)))
g<- g+geom_boxplot() 
g<- g+labs(title = "Tooth growth of 60 guinea pigs by dosage and delivery method of vitamin C", x = "Dose in milligrams/day", y = "Tooth Lengh")
g<- g+ scale_fill_discrete(name="dosage type")
g<- g+theme(plot.title = element_text(hjust = 0.5))
print(g)
```

## 4. Data Manipulation to separate data by dosage

```{r}
data_0.5<-subset(ToothGrowth, dose==0.5)
data_1<-subset(ToothGrowth, dose==1)
data_2<-subset(ToothGrowth, dose==2)
```

## 5. Performing T test to accept or reject hypothesis

```{r}
t1<-t.test(len~supp, data=data_0.5)
t2<-t.test(len~supp, data=data_1)
t3<-t.test(len~supp, data=data_2)
```

## 6. Creating DataFrame to analyze the result.

```{r}
Conclusion<-data.frame("p-value"=c(t1$p.value,t2$p.value,t3$p.value),
                       "Low-Confidence"=c(t1$conf.int[1],t2$conf.int[1],t3$conf.int[1]),
                       "High-Confidence"=c(t1$conf.int[2],t2$conf.int[2],t3$conf.int[2]),
                       "OJ mean" = c(t1$estimate[1],t2$estimate[1],t3$estimate[1]),
                       "VC mean" = c(t1$estimate[2],t2$estimate[2],t3$estimate[2]),
                       row.names = c("Dosage_.5","Dosage_1","Dosage_2"))

Conclusion
```

## 7.Conclusion

### 1. Since P value for dosage 0.5 and 1 are less than threshold value (0.5), we reject the null hypothesis which states that there is no relation between tooth growth and supplement method for 0.5 and 1mg dose. 

### 2. Since P value for dosage 2mg is greater than threshold value (0.5), we accept the null hypothesis which states that there is no relation between tooth growth and supplement method for 2mg dose.

### 3. From datafarme last 2 columns, it is concluded that OJ provides more growth than VC when dosage is 0.5mg or 1mg