---
title: "Statistical Inference (Simulaton Exercise)"
author: "Syed Shaheryar Tirmizi"
date: "7/27/2020"
---

## Overview
### In this exercise, I'm gonna perform 1000 simulations on 40 exponentials Data. Firstly, mean, variance and standard deviation of are calculated by general R manipulation. Then, by formula, mean, variance and standrd deviation is calculated and compare them. At the end, the data is visualized to check its approximation with Normal Distribution. 

## 1. Expected values Calculation of simulated data.

```{r}
library(ggplot2)
set.seed(555)
lamda<-0.2
n<-40
simulations<-1:1000
simulatedData<-data.frame(x=sapply(simulations, function(x) {mean(rexp(n,lamda))}))
meanData<-apply(simulatedData, 2, mean)
standarDeviataion<-apply(simulatedData, 2, sd)
Variance<-apply(simulatedData,2,var)
```
## 2. Expected values calculation by Formula of Center Limit Theorem.

```{r}
cltMean<-1/lamda
CLTstandardDeviation<-(1/lamda)/sqrt(n)
CLTvariance<-CLTstandardDeviation**2
```

## 3. Showing mean, variance and standard deviation of simulated data

```{r}
meanData
```
```{r}
Variance
```
```{r}
standarDeviataion
```

## 4. Showing mean, variance and standard deviation of Central Limit Theorem

```{r}
cltMean
```
```{r}
CLTvariance
```
```{r}
CLTstandardDeviation
```

## 5. Compare the expected values with the valus of central limit theorem.
### By showing values, it is concluded that expected values are approximately equal to the values of Central Limit Theorem.

## 6. Visualization of simulated data to approximate it with Normal Distribution

```{r}
g<-ggplot(simulatedData, aes(x=x))
g<-g+geom_histogram(aes(y=..density..),binwidth = 0.20)
g<-g+stat_function(fun=dnorm, args=list(mean=meanData, sd=standarDeviataion))
g<-g+geom_vline(xintercept = meanData, color="red")
print(g)
```
